---
title: "NCBI_SRA_file_download"
author: "Phil Morin"
date: "5/15/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Determine the Sequence Read Archive (SRA) files for a NCBI BioProject, and download all of them using the NCBI SRAtoolkit function "fastq-dump", 
This requires that SRAtoolkit is installed (see https://ncbi.github.io/sra-tools/install_config.html)

```{r}
# Bioproject ID
bioproject <- "SRP090345"

# path to SRAtoolkit directory
SRAtk_dir <- "~/Documents/Applications_local/sratoolkit.2.9.6-1-mac64/bin/"
# SRAtk_dir <- "C:\Users\phil.morin\Desktop\Programs_local\sratoolkit.2.9.6-1-win64_1\bin\"
```


# Get SRA Run Info for a BioProject using wget from Linux system:
wget 'http://trace.ncbi.nlm.nih.gov/Traces/sra/sra.cgi?save=efetch&rettype=runinfo&db=sra&term=PRJNA223640' -O - | tee SraRunInfo.csv
 This saves a .csv file with the SRP number (for multple runs in a project)
 use the SRP number below to get the SRA file info and ftp commands

```{r}
source('http://bioconductor.org/biocLite.R')
# biocLite("BiocUpgrade") # if upgrade message appears saying an upgrade is available.
#biocLite('SRAdb')
# or: # devtools::install_github("seandavi/SRAdb")
library(SRAdb)
#biocLite('DBI')
library(DBI)


######## only need to get SRA dbfile once; it's 2.5GB, so it takes a long time to load. After loaded, the file decompresses to ~32GB; copy the path to the saved file and set srafile = <path>.
# srafile_get = getSRAdbFile() 
      # write to file for future use
# write.csv(srafile_get, file="srafile_path.csv", row.names = FALSE, col.names = FALSE, quote = TRUE)
########

# set the path to the SRADBfile and extract the SRR list for the specified project. This uses the file "srafile_path.csv" saved above.
srafile = read.csv("srafile_path.txt", header = FALSE)
srafile1 <- as.character(srafile$V1)
con = dbConnect(RSQLite::SQLite(), srafile1)
SRRlist <- listSRAfile(bioproject, con)
write.csv(SRRlist, "NPho_SRR_list.csv")


# extract SRR column to create list of SRR files to get
SRRrun <- SRRlist$run

# or set up a short run based on specified list of SRR files
SRRshortlist <- c("SRR4292233", "SRR4292235")

```


# Run Fastq-dump from SRAtoolkit (from specified directory on computer, using bash)
SRAtoolkit commands in ~/Documents/Applications_local/sratoolkit.2.9.6-1-mac64/bin
```{r}

# create bash command to run 'fastq-dump' from SRAtoolkit directory
# test on short list of files.
#   fq_dump_srr <- paste0(SRAtk_dir,"fastq-dump --gzip --skip-technical --dumpbase --split-files --clip ", SRRshortlist)

# or get all files from the project
fq_dump_srr <- paste0(SRAtk_dir,"fastq-dump --gzip --skip-technical --dumpbase --split-files --clip ", SRRrun)

# run fastq-dump for each ERR. this splits the SRR file into R1 and R2 fastq file.
system(fq_dump_srr)

```

## Save data
```{r}
save.image(file=paste0("Bioproject_",bioproject,"_SRA_download.Rdata"))
```


